---
title: "Uganda Ebola Outbreak Report"
author: "James Azam & Marijn Hazelbag"
date: "13/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(incidence)
```
```{r echo=FALSE, message=FALSE}
ebola_dat <- read_csv(file = '../data/ebola_1.csv') %>% 
    arrange(reportDate)
```
## Report on the ongoing Ebola outbreak in Uganda, 2019/2020
The first Ebola virus disease (EVD) case was reported in Uganda on `r unclass(select(ebola_dat, reportDate)[1, ])$reportDate`. As of `r unclass(select(ebola_dat, reportDate)[nrow(ebola_dat), ])$reportDate`, there have been `r sum(table(ebola_dat$status))` reported cases in `r length(unique(ebola_dat$householdID))` unique households. Out of the total cases, `r as.numeric(table(ebola_dat$status)['confirmed'])` are confirmed, `r as.numeric(table(ebola_dat$status)['probable'])` probable, and `r as.numeric(table(ebola_dat$status)['suspected'])` suspected. The average number 
of cases per households reporting cases is `r mean(table(ebola_dat$householdID))`. 

The incidence curve below summarises the reported information.
```{r echo=FALSE, message=FALSE}
ebola_inc <- incidence(ebola_dat$onsetDate, groups = ebola_dat$status)
plot(ebola_inc, xlab = 'Date of onset')
```

