---
title: "Situation Report: Uganda Ebola Virus Disease Outbreak"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(incidence)
```
```{r echo=FALSE, message=FALSE}
ebola_dat <- read_csv(file = '../data/ebola_1_update.csv') %>% 
    arrange(reportDate) %>% 
  mutate(reportDelay = reportDate - onsetDate) 

ebola_dat_outbreak1 <- ebola_dat %>% filter(reportDate < reportDate[3])
ebola_dat_outbreak2 <- ebola_dat %>% filter(reportDate > reportDate[2])

```
```{r echo=FALSE, message=FALSE}

## overall calculations for report
first_reporting_date <- pull(ebola_dat, reportDate)[1]
last_reporting_date <- pull(ebola_dat, reportDate)[nrow(ebola_dat)]
total_confirmed <- as.numeric(table(ebola_dat$status)['confirmed'])
total_probable <- as.numeric(table(ebola_dat$status)['probable'])
total_deaths <- sum(!is.na(ebola_dat$deathDate))

#cases by sex

case_status_by_sex <- ebola_dat %>% 
    filter(status == 'confirmed') %>% 
    group_by(status, sex) %>% 
    count() 

##outbreak 1
total_confirmed_outbreak1 <- as.numeric(table(ebola_dat_outbreak1$status)['confirmed'])
total_probable_outbreak1  <- as.numeric(table(ebola_dat_outbreak1$status)['probable'])
total_suspected_outbreak1  <- as.numeric(table(ebola_dat_outbreak1$status)['suspected'])

##outbreak 2
total_confirmed_outbreak2 <- as.numeric(table(ebola_dat_outbreak2$status)['confirmed'])
total_probable_outbreak2  <- as.numeric(table(ebola_dat_outbreak2$status)['probable'])
total_suspected_outbreak2  <- as.numeric(table(ebola_dat_outbreak2$status)['suspected'])
n_deaths_outbreak2 <- sum(!is.na(ebola_dat_outbreak2$deathDate))


#case fatality ratio
#reportDelay_mean <- mean(ebola_dat_update$reportDelay)

fatality_confirmed_cases <- ebola_dat %>% 
    filter(status == 'confirmed') %>% 
    tidyr::drop_na(deathDate) %>% 
    nrow()

cfr <- binom.test(fatality_confirmed_cases, total_confirmed, conf.level = .95) %>% 
    broom::tidy(cfr) %>% 
    select(conf.low, estimate, conf.high)

## create the incidence object from the data
ebola_inc <- incidence(ebola_dat$onsetDate, groups = ebola_dat$status)
```
## External Situation Report 01: `r format(Sys.Date(),' %d-%m-%Y')`
<!-- The first Ebola virus disease (EVD) case was reported `r ebola_inc$timespan` days ago on `r first_reporting_date`.  -->

- From `r ebola_dat_outbreak1$reportDate[1]` to `r ebola_dat_outbreak1$reportDate[2]`, one probable and confirmed case were reported in Uganda but this outbreak did not take off.
- No deaths were reported from the outbreak during January, 2019.
- On `r ebola_dat_outbreak2$reportDate[3]`, a second outbreak was reported, and `r total_confirmed_outbreak2` total confirmed cases have been reported.
- `r n_deaths_outbreak2` number of deaths have been reported so far. 
 - As of `r last_reporting_date`, a total of `r nrow(ebola_dat)` EVD cases, including `r total_confirmed` confirmed and `r total_probable` probable cases have been reported in Uganda. 
 - Of the total confirmed cases, there have been `r fatality_confirmed_cases` deaths, with a case-fatality ratio of `r round(cfr$estimate*100, 1)`% (95% CI, `r round(cfr$conf.low *100, 1)`% to `r round(cfr$conf.high *100, 1)`%). 
- We suspect the case-fatality ratio is an underestimate of the actual value because we found a mean reporting delay of `r round(mean(ebola_dat$reportDelay))` days.
- Of the total confirmed, `r case_status_by_sex[1, 3]` are female and `r case_status_by_sex[2, 3]` are male.

- We have fitted a simple model to the reported data and it appears that we the peak of the outbreak has been reached and the outbreak is stable. A subgroup analysis suggests that there appears to be a decreasing trend in the number of cases in males and an increase in female cases. 
 
Below is an incidence curve by disease status of all the reported cases.
```{r echo=FALSE, message=FALSE}
plot(ebola_inc, xlab = 'Date of onset')
```

<!-- ## Additional data required -->
<!-- Like many other infectious diseases, Ebola virus disease (EVD) is spread by direct contact and over a spatial landscape. Hence, it will be useful for us to have GIS locations of households to ascertain social interactions and transmission chains. We will require demographic information(gender, age) for estimating metrics like the age-specific case-fatality ratios. These metrics are useful for recommending targetted interventions concerning age and sex, and to infer possible interactions. Additionally, we require information on hospitalisation (i.e. the date of hospitalisation, hospital name, and location). -->


